---
title:  'CMTH 642 - Assignment #3'
author: 'Submitted by: AQUINO, PETER'
output:
  word_document: default
---


Preparation:
The dataset is related to white Portuguese "Vinho Verde" wine. For more info: https://archive.ics.uci.edu/ml/datasets/Wine+Quality
Import to R the following file:
```{r}
source.file <- "http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv"

wine <- read.csv(file=source.file, header=TRUE, sep=";", na.strings=c("", "NA"), stringsAsFactors = FALSE)
```

##TASKS
#1. Check data characteristics. Is there missing data? Prepare the data for analysis.

```{r}
colSums(is.na(wine)) # there are no missing data
head(wine)
tail(wine)
str(wine)
summary(wine)
```

#2. What is the correlation between the attributes other than wine quality?

```{r}
#install.packages("corrplot")
library(corrplot)
wine.cor.matrix <- cor(wine[,-length(wine)])
wine.cor.matrix
corrplot(wine.cor.matrix, method="circle")
```



#Based on the above correlation matrix, the following attributes are highly correlated:
- residual.sugar / density
- density / alcohol

#, while the following attributes have a strong correlation:
- fixed.acidity / pH
- residual.sugar / alcohol
- free.sulfur.dioxide / total.sulfur.dioxide
- total.sulfur.dioxide / density 
- total.sulfur.dioxide / alcohol 


#Further analysis: Principal Component Analysis
```{r}
wine.pca <- princomp(wine[,-length(wine)], cor = TRUE, score = TRUE)
summary(wine.pca)
#eigen(cor(wine))
#eigen(cor(wine))$values/11
plot(wine.pca, type="lines", main = "Scree Plot for Wine Data")
biplot(wine.pca, main = "Biplot for Wine Data")
```
 
#3. Graph the frequency distribution of wine quality.

```{r}
hist(wine$quality, xlab = "Wine Quality", main = "Wine Quality Frequency Distribution", col = "grey")
```

#4. Reduce the levels of rating for quality to three levels as high, medium and low.

```{r}
#Qualty levels: low=3-5, medium=6-7, high=8-9
#Following creates another column to store these 3 quality levels
wine$qlevel <- cut(wine$quality, breaks = 3, labels = c("low","medium","high"))
```

#5. Normalize the data set.

```{r}
#The following min-max normalization function will be applied to all attributes except quality and qlevel

norm.attrib <- function(target.attrib) {
    return ((target.attrib - min(target.attrib)) / (max(target.attrib) - min(target.attrib)))
}

wine.norm <- as.data.frame(lapply(wine[1:11], norm.attrib))

# Re-instate quality levels to the normalized dataset
wine.norm$qlevel <- as.factor(wine$qlevel)
```

#6. Divide the data to training and testing groups.

```{r}
library(caret)
#set.seed(888)
train.set.index <- createDataPartition(wine.norm$qlevel, p=0.70, list=FALSE)
train.set <- wine.norm[train.set.index, ]
test.set <- wine.norm[-train.set.index, ]
```

#7. Use the kNN algorithm to predict the quality of wine using its attributes.

```{r}
#Evaluate different k values
train.ctl <- trainControl(method="repeatedcv", repeats=3)
kValues <- train(qlevel~., data = train.set, method="knn", trControl=train.ctl, tuneLength=10)
kValues

#Get the optimal k to be used in the kNN algorithm
kresults <- as.list(kValues)
kresults <- as.data.frame(kresults$results)
kbest <- kresults[which.max(kresults$Accuracy),c('k')]
kbest

#Run kNN algorithm with k = kbest
library(class)
wine.knn <- knn(train.set[, -12], test.set[, -12], cl = train.set$qlevel, k = kbest)
```

#8. Evaluate the model performance.

```{r}
#install.packages('e1071')
library(e1071)
confusionMatrix(test.set$qlevel,wine.knn)
```

# Alternatively, run kNN for each k value and display each model performance
```{r}
k.values <- c(kresults[, 1])

for (kval in k.values) { 
  print(paste('>>>>> k = ', kval))
  wine.knn <- knn(train.set[, -12], test.set[, -12], cl = train.set$qlevel, k = kval)
  print(confusionMatrix(test.set$qlevel,wine.knn))
  print('<<<<<<<<<<<<<<<<<<<')
}
```


END of Assignment #3.
